---
title: "Non-Vote Rates for PHL Judicial Elections (2015-2019)"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r setup, include=FALSE}
library(flexdashboard)
library(readxl)
library(ggplot2)
library(dplyr)
library(plotly)
library(rjson)
library(DT)


##import dataset

Judge <- read_xlsx("myworkbook.xlsx", sheet="Total")

Ward <- read_xlsx("myworkbook.xlsx", sheet = "Ward")

##round NON_VOTE_RATE COLUMN

Judge$VOTERS <- round(Judge$VOTERS, digits = 0) 
Judge$NON_VOTE_RATE <- round(Judge$NON_VOTE_RATE, digits = 2)

##round NON_VOTE_RATE COLUMN Ward

Ward$VOTERS <- round(Ward$VOTERS, digits = 0) 
Ward$NON_VOTE_RATE <- round(Ward$NON_VOTE_RATE, digits = 2)

AVG_WARD <- Ward %>% filter(CATEGORY != "MAYOR" & CATEGORY != "DISTRICT ATTORNEY") %>%
  group_by(WARD) %>% summarise("NON_VOTE_RATE" = round(mean(NON_VOTE_RATE),2))

```

# Maps

Column {data-width=750}{.tabset}
-----------------------------------------------------------------------

### All Courts

```{r}
## Choropleth Map
## Use AVG_WARD Data

phl_geo = "Political_Wards.geojson"



geojson <- rjson::fromJSON(file=phl_geo)
df <- AVG_WARD
g <- list(
  fitbounds = "locations",
  visible = FALSE
)
fig <- plot_ly() 
fig <- fig %>% add_trace(
  type="choropleth",
  geojson=geojson,
  locations=df$WARD,
  z=df$NON_VOTE_RATE,
  colorscale="Viridis",
  featureidkey="properties.WARD_NUM"
)
fig <- fig %>% layout(
  geo = g
)
fig <- fig %>% colorbar(title = "Non-Vote Rate")
fig <- fig %>% layout(
  title = "Avg Non-Vote Rate by Ward")

fig


```

### Commonwealth Court

```{r}
## Commonwealth court by ward

AVG_WARD2 <- Ward %>% filter(CATEGORY == "COMMONWEALTH COURT") %>%
  group_by(WARD) %>% summarise("NON_VOTE_RATE" = mean(NON_VOTE_RATE))


phl_geo = "Political_Wards.geojson"



geojson <- rjson::fromJSON(file=phl_geo)
df <- AVG_WARD2
g <- list(
  fitbounds = "locations",
  visible = FALSE
)
fig1 <- plot_ly() 
fig1 <- fig1 %>% add_trace(
  type="choropleth",
  geojson=geojson,
  locations=df$WARD,
  z=df$NON_VOTE_RATE,
  colorscale="Viridis",
  featureidkey="properties.WARD_NUM"
)
fig1 <- fig1 %>% layout(
  geo = g
)
fig1 <- fig1 %>% colorbar(title = "Non-Vote Rate")
fig1 <- fig1 %>% layout(
  title = "Avg Non-Vote Rate by Ward (Commonwealth Court)")

fig1
```

### Common Pleas Court

```{r}
## Common pleas court by ward

AVG_WARD3 <- Ward %>% filter(CATEGORY == "COURT OF COMMON PLEAS") %>%
  group_by(WARD) %>% summarise("NON_VOTE_RATE" = mean(NON_VOTE_RATE))


phl_geo = "Political_Wards.geojson"



geojson <- rjson::fromJSON(file=phl_geo)
df <- AVG_WARD3
g <- list(
  fitbounds = "locations",
  visible = FALSE
)
fig2 <- plot_ly() 
fig2 <- fig2 %>% add_trace(
  type="choropleth",
  geojson=geojson,
  locations=df$WARD,
  z=df$NON_VOTE_RATE,
  colorscale="Viridis",
  featureidkey="properties.WARD_NUM"
)
fig2 <- fig2 %>% layout(
  geo = g
)
fig2 <- fig2 %>% colorbar(title = "Non-Vote Rate")
fig2 <- fig2 %>% layout(
  title = "Avg Non-Vote Rate by Ward (Court of Common Pleas)")

fig2
```

### Municipal Court

```{r}
## Municipal Court by Ward

AVG_WARD4 <- Ward %>% filter(CATEGORY == "MUNICIPAL COURT") %>%
  group_by(WARD) %>% summarise("NON_VOTE_RATE" = mean(NON_VOTE_RATE))


phl_geo = "Political_Wards.geojson"



geojson <- rjson::fromJSON(file=phl_geo)
df <- AVG_WARD4
g <- list(
  fitbounds = "locations",
  visible = FALSE
)
fig3 <- plot_ly() 
fig3 <- fig3 %>% add_trace(
  type="choropleth",
  geojson=geojson,
  locations=df$WARD,
  z=df$NON_VOTE_RATE,
  colorscale="Viridis",
  featureidkey="properties.WARD_NUM"
)
fig3 <- fig3 %>% layout(
  geo = g
)
fig3 <- fig3 %>% colorbar(title = "Non-Vote Rate")
fig3 <- fig3 %>% layout(
  title = "Avg Non-Vote Rate by Ward (Municipal Court)")

fig3

```

### PA Supreme Court

```{r}
## PA Supreme Court by Ward

AVG_WARD5 <- Ward %>% filter(CATEGORY == "PA SUPREME COURT") %>%
  group_by(WARD) %>% summarise("NON_VOTE_RATE" = mean(NON_VOTE_RATE))


phl_geo = "Political_Wards.geojson"



geojson <- rjson::fromJSON(file=phl_geo)
df <- AVG_WARD5
g <- list(
  fitbounds = "locations",
  visible = FALSE
)
fig4 <- plot_ly() 
fig4 <- fig4 %>% add_trace(
  type="choropleth",
  geojson=geojson,
  locations=df$WARD,
  z=df$NON_VOTE_RATE,
  colorscale="Viridis",
  featureidkey="properties.WARD_NUM"
)
fig4 <- fig4 %>% layout(
  geo = g
)
fig4 <- fig4 %>% colorbar(title = "Non-Vote Rate")
fig4 <- fig4 %>% layout(
  title = "Avg Non-Vote Rate by Ward (PA Supreme Court)")

fig4

```

### Superior Court

```{r}
## Superior Court by Ward

AVG_WARD6 <- Ward %>% filter(CATEGORY == "SUPERIOR COURT") %>%
  group_by(WARD) %>% summarise("NON_VOTE_RATE" = mean(NON_VOTE_RATE))


phl_geo = "Political_Wards.geojson"



geojson <- rjson::fromJSON(file=phl_geo)
df <- AVG_WARD6
g <- list(
  fitbounds = "locations",
  visible = FALSE
)
fig5 <- plot_ly() 
fig5 <- fig5 %>% add_trace(
  type="choropleth",
  geojson=geojson,
  locations=df$WARD,
  z=df$NON_VOTE_RATE,
  colorscale="Viridis",
  featureidkey="properties.WARD_NUM"
)
fig5 <- fig5 %>% layout(
  geo = g
)
fig5 <- fig5 %>% colorbar(title = "Non-Vote Rate")
fig5 <- fig5 %>% layout(
  title = "Avg Non-Vote Rate by Ward (Superior Court)")

fig5
```

Column {data-width=250}
-----------------------------------------------------------------------

### Info

#### Data Source and Goals

Original datasets provided by [The Office of the Philadelphia commisioners](ttps://www.philadelphiavotes.com/en/resources-a-data/archived-data-sets)

This project looked at the last three Philadelphia Elections that had Judicial Races. The goal was to determine the rate at which voters, who did vote in these elections, do not vote for Judges (Non-Vote Rate).

- [2019 General Election](https://phillyvotes-files.s3.amazonaws.com/raw-data/2019_general.csv)
- [2017 General Election](https://files7.philadelphiavotes.com/election-results/2017_GENERAL/2017_GENERAL_-_ELECTION_OFFICES.xlsx#_ga=2.224334683.1734944450.1612730933-1470740827.1611798905)
- [2015 General Election](https://files7.philadelphiavotes.com/election-results/2015_GENERAL/2015_GENERAL_-_WARD_RESULTS_-_PUBLIC.xlsx#_ga=2.223108955.1734944450.1612730933-1470740827.1611798905)

These elections include Judicial Races for:

- Commonwealth Court
- Court of Common Pleas
- Municipal Court
- PA Supreme Court
- Superior Court

#### Methodology

There is no metric for Non-Vote Rate in the original datasets, however, a reasonably accurate Non-Vote Rate was attained by comparing the number of votes for the top of the ballot race against the votes cast for the respective Judicial races.



# Graphs


Column {data-width=750}{.tabset}
-----------------------------------------------------------------------

### by Ward

```{r}
## Total Average NON_VOTE_RATE by WARD

ward_overview <- ggplot(AVG_WARD, aes(x = WARD, y = NON_VOTE_RATE)) +
  geom_bar(stat="identity") +
  ylab("Non-Vote Rate") +
  xlab("Ward #")

ggplotly(ward_overview)
```

### by Ward and Court

```{r}
## NON_VOTE_RATE by WARD and CATEGORY

AVG_WARD1 <- Ward %>% filter(CATEGORY != "MAYOR" & CATEGORY != "DISTRICT ATTORNEY") %>%
  group_by(WARD, CATEGORY) %>% summarise("NON_VOTE_RATE" = mean(NON_VOTE_RATE))

ward_overview1 <- ggplot(AVG_WARD1, aes(x = WARD, y = NON_VOTE_RATE)) +
  geom_bar(stat="identity", width = .5) +
  ylab("Non-Vote Rate") +
  xlab("Ward #") +
  facet_wrap("CATEGORY")

ggplotly(ward_overview1)
```

Column {data-width=250}{.tabset}
-----------------------------------------------------------------------

### All Elections

```{r}
## Total NON_VOTE_RATE

NON_VOTE_TOTALS <- Judge %>% filter(CATEGORY != "MAYOR" & CATEGORY != "DISTRICT ATTORNEY") %>%
  group_by(CATEGORY) %>% summarise("NON_VOTE_RATE" = round(mean(NON_VOTE_RATE),2))

non_vote_total <- ggplot(NON_VOTE_TOTALS, aes(x = CATEGORY, y = NON_VOTE_RATE)) +
  geom_bar(stat="identity") +
  ylab("Non-Vote Rate") +
  xlab("") + 
  theme(axis.text.x = element_text(size=7))

ggplotly(non_vote_total)
```

### 2019

```{r}
Judge_2019 <- Judge %>% filter(YEAR == 2019 & CATEGORY != "MAYOR")

p2019 <- ggplot(Judge_2019, aes(x = CATEGORY, y = NON_VOTE_RATE)) +
  geom_bar(stat="identity") +
  ylab("Non-Vote Rate") +
  xlab("") +
  ggtitle("2019 General Election")

ggplotly(p2019)
```

### 2017

```{r}
Judge_2017 <- Judge %>% filter(YEAR == 2017 & CATEGORY != "DISTRICT ATTORNEY")

p2017 <- ggplot(Judge_2017, aes(x = CATEGORY, y = NON_VOTE_RATE)) +
  geom_bar(stat="identity") +
  ylab("Non-Vote Rate") +
  xlab("") +
  ggtitle("2017 General Election") + 
  theme(axis.text.x = element_text(size=7))

ggplotly(p2017)
```

### 2015

```{r}
Judge_2015 <- Judge %>% filter(YEAR == 2015 & CATEGORY != "MAYOR")

p2015 <- ggplot(Judge_2015, aes(x = CATEGORY, y = NON_VOTE_RATE)) +
  geom_bar(stat="identity") +
  ylab("Non-Vote Rate") +
  xlab("") +
  ggtitle("2015 General Election") + 
  theme(axis.text.x = element_text(size=7))

ggplotly(p2015)
```

# Data

Column {data-width=1000}{.tabset}
-----------------------------------------------------------------------

### Total

```{r}
datatable(Judge,
          extensions = c('Buttons', 'Scroller'),
          options = list(
            dom = 'Bfrtip',
            buttons = c('copy', 'print', 'csv'),
            scrollX=TRUE)
          )
          
```

### By Ward

```{r}
datatable(AVG_WARD,
          extensions = c('Buttons', 'Scroller'),
          options = list(
            dom = 'Bfrtip',
            buttons = c('copy', 'print', 'csv'),
            scrollX=TRUE)
          )
          
```
